\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-07-08 scratch/notes_john/JSS-article/JSS-article-3.tex
some changes to John's JSS draft

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-3_files/figure-latex/mpg-horsepower-scatterplot-polynomials-1}

@@

\caption{\code{mpg} vs. \code{horsepower} for the \code{Auto} data, showing fitted polynomials of degree 1 through 5.}\label{fig:mpg-horsepower-scatterplot-polynomials}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics[width=0.5\linewidth]{JSS-article-3_files/figure-latex/mpg-horsepower-MSE-se-1}

\caption[Estimated squared error as a function of polynomial degree, $p$]{Estimated squared error as a function of polynomial degree, $p$}\label{fig:mpg-horsepower-MSE-se}
\end{figure}
\end{CodeChunk}

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 431467
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.192
bias-adjusted cross-validation criterion = 19.18107
95% CI for bias-adjusted CV criterion = (15.71858, 22.64356)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> Auto$cylinders <- factor(Auto$cylinders,
+                          labels=c("3-4", "3-4", "5-6", "5-6", "8"))
R> Auto$year <- as.factor(Auto$year)
R> Auto$origin <- factor(Auto$origin,
+                       labels=c("America", "Europe", "Japan"))
R> rownames(Auto) <- make.names(Auto$name, unique=TRUE)
R> Auto$name <- NULL
R> 
R> scatterplotMatrix(~ mpg + displacement + horsepower + weight + acceleration, 
+                   smooth=list(spread=FALSE), data=Auto, pch=".")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-3_files/figure-latex/Auto-explore-1}

@@

\caption[Scatterplot matrix for the numeric variables in the \code{Auto} data]{Scatterplot matrix for the numeric variables in the \code{Auto} data. In each panel, the solid line shows the linear least-squares fit and the broken line is for a nonparametric regression.}\label{fig:Auto-explore}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto <- lm(mpg ~ ., data = Auto)
R> crPlots(m.auto, pch=".")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-3_files/figure-latex/Auto-working-model-1}

\caption[Component+residual plots for the working model fit to the \code{Auto} data]{Component+residual plots for the working model fit to the \code{Auto} data. The broken blue lines show linear least-squares fits, while the solid magenta lines are nonparametric-regression smooths.}\label{fig:Auto-working-model}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
+                   + acceleration, 
+                   smooth=list(spread=FALSE), data=A, pch=".")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-3_files/figure-latex/Auto-transformed-scatterplot-matrix-1}

@@

\caption[Scatterplot matrix for the transformed numeric variables in the \code{Auto} data]{Scatterplot matrix for the transformed numeric variables in the \code{Auto} data}\label{fig:Auto-transformed-scatterplot-matrix}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m, pch=".")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{JSS-article-3_files/figure-latex/Auto-CR-plots-transformed-1}

\caption[Component+residual plots for the model fit to the transformed \code{Auto} data]{Component+residual plots for the model fit to the transformed \code{Auto} data}\label{fig:Auto-CR-plots-transformed}
\end{figure}
\end{CodeChunk}

@@

\subsection[Example: Applying recursive CV to polynomial regression for
the Auto data]{\texorpdfstring{Example: Applying recursive CV to
polynomial regression for the Auto
data\footnote{What we call here ``recursive CV'' has also been termed
  ``nested CV'' \citep[e.g., in the \pkg{cvms} package
  by][]{OlsenZachariae:2024}. We prefer, however, to reserve the term
  ``nested CV'' for the procedure described by
  \citet{BatesHastieTibshirani:2023}.}}{Example: Applying recursive CV to polynomial regression for the Auto data}}\label{example-applying-recursive-cv-to-polynomial-regression-for-the-auto-datarecursive-cv}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13fb26ea8>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

>>>>

{\centering \subfloat[\label{fig:mpg-horsepower-polynomials-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/mpg-horsepower-polynomials-1} }\subfloat[\label{fig:mpg-horsepower-polynomials-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/mpg-horsepower-polynomials-2} }

@@

\caption{(a) \code{mpg} vs. \code{horsepower} for the \code{Auto} data, showing fitted polynomials of degree 1 through 5. (b) Estimated squared error as a function of polynomial degree, $p$.}\label{fig:mpg-horsepower-polynomials}
\end{figure}
\end{CodeChunk}

@@

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 180983
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.37365
bias-adjusted cross-validation criterion = 19.35319
95% CI for bias-adjusted CV criterion = (15.87805, 22.82834)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> Auto$cylinders <- factor(Auto$cylinders,
+                          labels=c("3-4", "3-4", "5-6", "5-6", "8"))
R> Auto$year <- as.factor(Auto$year)
R> Auto$origin <- factor(Auto$origin,
+                       labels=c("America", "Europe", "Japan"))
R> rownames(Auto) <- make.names(Auto$name, unique=TRUE)
R> Auto$name <- NULL
R> 
R> scatterplotMatrix(~ mpg + displacement + horsepower + weight + acceleration, 
+                   smooth=list(spread=FALSE), data=Auto, pch=".")
R> 
R> m.auto <- lm(mpg ~ ., data = Auto)
R> crPlots(m.auto, pch=".")
\end{CodeInput}
\begin{figure}

{\centering \subfloat[\label{fig:Auto-explore-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/Auto-explore-1} }\subfloat[\label{fig:Auto-explore-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/Auto-explore-2} }

@@

\caption{(a) Scatterplot matrix for the numeric variables in the \code{Auto} data. In each panel, the solid line shows the linear least-squares fit and the broken line is for a nonparametric regression. (b) Component+residual plots for the working model fit to the \code{Auto} data. The broken blue lines show linear least-squares fits, while the solid magenta lines are nonparametric-regression smooths.}\label{fig:Auto-explore}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> Auto$cylinders <- factor(Auto$cylinders,
+                          labels=c("3-4", "3-4", "5-6", "5-6", "8"))
R> Auto$year <- as.factor(Auto$year)
R> Auto$origin <- factor(Auto$origin,
+                       labels=c("America", "Europe", "Japan"))
R> rownames(Auto) <- make.names(Auto$name, unique=TRUE)
R> Auto$name <- NULL
R> 
R> scatterplotMatrix(~ mpg + displacement + horsepower + weight + acceleration, 
+                   smooth=list(spread=FALSE), data=Auto, pch=".")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto <- lm(mpg ~ ., data = Auto)
R> crPlots(m.auto, pch=".")
\end{CodeInput}
\end{CodeChunk}

@@

{\centering \subfloat[\label{fig:Auto-transformed-scatterplot-matrix-1}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/Auto-transformed-scatterplot-matrix-1} }\subfloat[\label{fig:Auto-transformed-scatterplot-matrix-2}]{\includegraphics[width=0.45\linewidth]{JSS-article-3_files/figure-latex/Auto-transformed-scatterplot-matrix-2} }

@@

\caption{(a) Scatterplot matrix for the transformed numeric variables in the \code{Auto} data. (b) Component+residual plots for the working model fit to the transformed \code{Auto} data.}\label{fig:Auto-transformed-scatterplot-matrix}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
+                   + acceleration, 
+                   smooth=list(spread=FALSE), data=A, pch=".")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> crPlots(m, pch=".")
\end{CodeInput}
\end{CodeChunk}

@@

\subsection[Example: Applying recursive CV to polynomial regression for
the Auto data]{\texorpdfstring{Example: Applying recursive CV to
polynomial regression for the Auto
data\footnote{What we call here ``recursive CV'' has also been termed
  ``nested CV'' \citep[e.g., in the \pkg{cvms} package
  by][]{OlsenZachariae:2024}. We prefer, however, to reserve the term
  ``nested CV'' for the procedure described by
  \citet{BatesHastieTibshirani:2023}.}}{Example: Applying recursive CV to polynomial regression for the Auto data}}\label{example-applying-recursive-cv-to-polynomial-regression-for-the-auto-data9}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x10f0f2c50>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}
